---
title: "Federal Employees"
output: 
  html_document:
    toc: true
    code_folding: hide
---

```{r setup, include=FALSE}
source("setup.R")
  load("FACTDATA_DEC2016.Rdata")
  data <- d
  
  # names of things
data %<>% 
  left_join(read.csv("opm/DTagy.csv"), by = "AGYSUB") %>%
  left_join(read.csv("opm/DTtoa.csv"), by = "TOA") %>%
  left_join(read.csv("opm/DTedlvl.csv"), by = "EDLVL") %>%
  left_join(read.csv("opm/DTocc.csv"), by = "OCC") %>%
  left_join(read.csv("opm/DTloc.csv"), by = "LOC") 
```
 

# The Department of Transportation
```{r}
# subset to DOT
data %<>% filter(AGY == "TD") # DOT only
```

```{r all-employees} 
# all employees
ggplot(data) + geom_bar(aes(x=AGYSUBT)) + coord_flip() + 
  labs(x = "", y = "", title = "Total Employees")
```

---

# GS levels

FAA uses different system.
```{r ses-gs, fig.height=5}
# by GS level (note FAA has a GS grade even for these employees)
ggplot(data) + 
  geom_bar(aes(x=AGYSUBT, fill = GSEGRD)) +
  coord_flip() + 
  labs(x = "", y = "", title = "")
```

---

#FAA

```{r FAA}
ggplot(data %>% filter(
  LOC == "11",
  AGYSUBT == "TD03-FEDERAL AVIATION ADMINISTRATION",
  #TOAT == "48-Excepted Service - Other",
  grepl("EX|FJ|EV", PPGRD)
) ) + coord_flip() + 
  geom_bar(aes(OCCT, fill = PPGRD)) + 
  labs(x = "", y = "", title = "FAA \"Exempted Service\" non-Air Traffic Control Employees by pay grade")  +
  theme(legend.position="bottom", legend.title = element_blank())

data %<>% 
  mutate(AGYSUBT = str_remove(AGYSUBT, "^TD...") ) %>% 
  mutate(AGYSUBT = str_to_title(AGYSUBT))
```

We recode FAA employees: Appointees are those starting with an EX or FJ (Ex is executive, and FJ is non-career appointments).  EV 1-4 are different classes of SES employees.  The FVs are for most employees (closest to the GS scale).  FG is for employees with specialized skills such as air traffic controllers.
```{r recode-FAA, echo=TRUE}
# recode FAA appointees: appointees are those starting with an EX or FJ (Ex is executive, and FJ is non-career appointments).  EV 1-4 are different classes of SES employees.  The FVs are for most employees (closest to the GS scale).  FG is for employees with specialized skills such as air traffic controllers.
data %<>% 
  mutate(appointment = TOAT) %>% 
  mutate(appointment = ifelse(
    str_detect(PPGRD, "FJ"),
    "44-Excepted Service - Schedule C",
    appointment)) %>%
  mutate(appointment = ifelse(
    str_detect(PPGRD, "EV"),
    "50-Senior Executive Service - Career",
    appointment)) %>%
  mutate(appointment = ifelse(
    str_detect(PPGRD, "EX"),
    "Administrator",
    appointment)) 

# Identify SES political appointees
data %<>% 
  mutate(
    political = ifelse(appointment %in% c("44-Excepted Service - Schedule C", 
                                   "55-Senior Executive Service - Non-Career", 
                                   "60-Senior Executive Service - Limited Term",
                                   "65-Senior Executive Service - Limited Emergency",
                                   "Administrator"), 
                       "Political", NA)) %>%
  # Carreer SES:
  mutate(political = ifelse(appointment %in% c("50-Senior Executive Service - Career"), 
                       "Carreer", political))
```


---

# Senior Executive Service Employees

```{r ses-employees} 
# SES employees 
data %>% 
  drop_na(political) %>% 
ggplot() + 
  geom_bar(aes(x=AGYSUBT, fill = appointment)) +
  coord_flip() + 
  labs(x = "", y = "", 
       fill= "",
       title = "SES Employees")
```

---
# Pay grade: 

```{r ses-by-pay}
# by GS level (note FAA has lots of non-GS)
ggplot(data %>% filter(grepl("GS|EX|FJ|EV", PPGRD))) + 
  geom_bar(aes(x=AGYSUBT, fill = PPGRD)) +
  coord_flip() + 
  labs(x = "", y = "", 
       fill = "Pay Grade",
       title = "SES by Pay Grade")
```


---

Some of these occupations are more likely to be involved in rule writing:
```{r ses-occupations, fig.height=5.5}
data %>% filter(#LOC == "11",
                appointment == "50-Senior Executive Service - Career")%>% 
ggplot()+
  coord_flip() + 
  geom_bar(aes(AGYSUBT, fill = OCCT)) + 
  labs(x = "", y = "", 
       fill = "Occupation",
       title = "SES Employees")  + 
  guides(fill = guide_legend(ncol = 1))

```

---


# Political v. Career SES

```{r political}

ggplot(data %>% filter(!is.na(political), LOC == "11")) +
  geom_bar(aes(x=AGYSUBT, fill = political)) +
  coord_flip() + 
  labs(x = "", y = "Number of Employees", 
       title = "Senior-Level Employees in DC")  +
  theme(legend.position="bottom", legend.title = element_blank())

data$SALARY %<>% as.numeric()

data %>% 
  filter(!is.na(political), LOC == "11") %>% 
  group_by(AGYSUBT, political) %>% 
  summarise(budget = sum(SALARY)) %>% 
ggplot() +
  geom_col(aes(x=AGYSUBT, y = budget, fill = political)) +
  coord_flip() + 
  scale_y_continuous(labels = scales::dollar) +
  labs(x = "", y = "", title = "Senior-Level Salaries in DC")  +
  theme(legend.position="bottom", legend.title = element_blank())

data %>% 
  filter(AGYSUB %in% c("TD15", "TD16", "TD10","TD05", "TD17", "TD04", "TD03" )) %>% 
  filter(!is.na(political), LOC == "11") %>% 
  group_by(AGYSUBT, political) %>% 
  summarise(budget = sum(SALARY)) %>%
ggplot() +
  geom_col(aes(x=AGYSUBT, y = budget, fill = political)) +
  coord_flip() + 
  scale_y_continuous(labels = scales::dollar) +
  labs(x = "", y = "", title = "Senior-Level Salaries in DC of Agencies in Regulatory Budget")  +
  theme(legend.position="bottom", legend.title = element_blank())
```

```{r}
save(data, file = "data/FederalEmployees.Rdata")
```

# Regulatory Budget

```{r reg-budget-data}
b <- readxl::read_xlsx(here("data/RegulatorsBudget_RawData-2019.xlsx"))
b %<>% select(Agency, '2016') %>% 
  filter(Agency %in% c("Federal Aviation Administration (18)",
                       "Federal Highway Administration (19)",
                       "Federal Motor Carrier Safety Administration (20)",
                       "Federal Railroad Administration",
                       "National Highway Traffic Safety Administration",
                       "Pipeline and Hazardous Materials Safety Administration (21)",
                       "Surface Transportation Board (22)") )

b %<>% rename(Budget = `2016`)%>% 
  mutate(Agency = str_remove(Agency, " .[0-9][0-9]."))

b$Budget %<>% as.numeric()

b %<>% mutate(Budget = Budget*1000000)
```

```{r reg-budget}
b %>% 
ggplot() +
  geom_col(aes(x=Agency, y = Budget)) +
  coord_flip() + 
  scale_y_continuous(labels = scales::dollar) +
  labs(x = "", y = "", title = "Regultory budget")  +
  theme(legend.position="bottom", legend.title = element_blank())
```


```{r}
RegBudget <- b
save(RegBudget, file = "data/RegBudget.Rdata")
```